---
title: "Geospatial plots"
output: html_notebook
---

Code based on example from: https://www.r-spatial.org/r/2018/10/25/ggplot2-sf-2.html

```{r}
library("ggplot2")
library("sf")
library("rnaturalearth")
library("rnaturalearthdata")
#library("rnaturalearthhires")    # For hi-res country maps (scale="large")
library("ggspatial")
library("maps")
library("ggrepel")
library("cowplot")                # For multiple subplots
```


# Geospatial plot

```{r}
# Data
world <- ne_countries(scale = "medium", returnclass = "sf")
world_points <- cbind(world, st_coordinates(st_centroid(world$geometry)))

# Geospatial plot
ggplot(data = world) + 
  geom_sf(fill = "antiquewhite") + 
  geom_text(data = world_points, aes(x=X, y=Y, label=name), color = "darkblue", fontface = "bold", check_overlap = FALSE) + 
  annotate(geom = "text", x = -90, y = 26, label = "Gulf of Mexico", fontface = "italic", color = "grey22", size = 6) + 
  annotation_scale(location = "bl", width_hint = 0.5) + 
  annotation_north_arrow(location = "bl", which_north = "true", pad_x = unit(1.4, "cm"), pad_y = unit(.8, "cm"), style = north_arrow_fancy_orienteering) + 
  coord_sf(xlim = c(-102.15, -74.12), ylim = c(7.65, 33.97), expand = FALSE) + 
  xlab("Longitude") + 
  ylab("Latitude") + 
  ggtitle("Map of the Gulf of Mexico and the Caribbean Sea") + 
  theme(panel.grid.major = element_line(color = gray(.5), linetype = "dashed", size = 0.5), panel.background = element_rect(fill = "aliceblue"))
```


# Adding layers

```{r}
# More data
states <- st_as_sf(map("state", plot = FALSE, fill = TRUE))
sites <- data.frame(longitude = c(-80.144005, -80.109), latitude = c(26.479005, 26.83))
sites <- st_as_sf(sites, coords = c("longitude", "latitude"), crs = 4326, agr = "constant")
states <- cbind(states, st_coordinates(st_centroid(states)))
states$nudge_y <- -1
states$nudge_y[states$ID == "Florida"] <- 0.5
states$nudge_y[states$ID == "South Carolina"] <- -1.5
counties <- st_as_sf(map("county", plot = FALSE, fill = TRUE))
counties <- subset(counties, grepl("florida", counties$ID))
counties$area <- as.numeric(st_area(counties))
flcities <- data.frame(state = rep("Florida", 5), city = c("Miami", "Tampa", "Orlando", "Jacksonville", "Sarasota"), 
                       lat = c(25.7616798, 27.950575, 28.5383355, 30.3321838, 27.3364347), 
                       lng = c(-80.1917902, -82.4571776, -81.3792365, -81.655651, -82.5306527))
flcities <- st_as_sf(flcities, coords = c("lng", "lat"), remove = FALSE,crs = 4326, agr = "constant")

# Detailed plot
ggplot(data = world) +
    geom_sf(fill = "antiquewhite1") +
    geom_sf(data = states, fill = NA) + 
    geom_sf(data = counties, aes(fill = area)) +
    geom_sf(data = sites, size = 4, shape = 23, fill = "darkred") +
    geom_sf(data = flcities) +
    geom_text_repel(data = flcities, aes(x = lng, y = lat, label = city), fontface = "bold", 
                    nudge_x = c(1, -1.5, 2, 2, -1), nudge_y = c(0.25, -0.25, 0.5, 0.5, -0.5)) +
    geom_label(data = states, aes(X, Y, label = ID), size = 5, fontface = "bold", nudge_y = states$nudge_y) +
    scale_fill_viridis_c(trans = "sqrt", alpha = .4) +
    annotation_scale(location = "bl", width_hint = 0.4) +
    annotation_north_arrow(location = "bl", which_north = "true", 
        pad_x = unit(0.5, "in"), pad_y = unit(0.4, "in"),
        style = north_arrow_fancy_orienteering) +
    coord_sf(xlim = c(-88, -78), ylim = c(24.5, 33), expand = FALSE) +
    xlab("Longitude") + ylab("Latitude") +
    ggtitle("Observation Sites", subtitle = "(2 sites in Palm Beach County, Florida)") +
    theme(panel.grid.major = element_line(color = gray(0.5), linetype = "dashed", 
        size = 0.5), panel.background = element_rect(fill = "aliceblue"))
```



# Multiple plots using Cowplot

```{r fig.width=16}
gworld <- {ggplot(data = world) +
  geom_sf(aes(fill = region_wb)) +
  geom_rect(xmin = -102.15, xmax = -74.12, ymin = 7.65, ymax = 33.97, 
     fill = NA, colour = "black", size = 1.5) +
  scale_fill_viridis_d(option = "plasma") +
  theme(panel.background = element_rect(fill = "azure"),
     panel.border = element_rect(fill = NA))}

ggulf <- {ggplot(data = world) +
  geom_sf(aes(fill = region_wb)) +
  annotate(geom = "text", x = -90, y = 26, label = "Gulf of Mexico", 
     fontface = "italic", color = "grey22", size = 3) +
  coord_sf(xlim = c(-102.15, -74.12), ylim = c(7.65, 33.97), expand = FALSE) +
  scale_fill_viridis_d(option = "plasma") +
  theme(legend.position = "none", axis.title.x = element_blank(), 
     axis.title.y = element_blank(), panel.background = element_rect(fill = "azure"), 
     panel.border = element_rect(fill = NA))}

plot_grid(gworld, ggulf, nrow = 1, rel_widths = c(2.3, 1))
```




# sf package (simple features)

https://cran.r-project.org/web/packages/sf/vignettes/sf1.html


```{r}
library(sf)
nc <- st_read(system.file("shape/nc.shp", package="sf"), quiet = TRUE)
#st_write(nc, "nc.shp", delete_layer = TRUE)     # delete_layer = TRUE overwrites existing layer
```

```{r}
class(nc)
```


```{r}
methods(class = 'sf')
```


```{r}
plot(nc)         # Standard plotting method (geom_sf in ggplot2 is recommended)
```



```{r}
nc_geom <- st_geometry(nc)
nc_geom
```

```{r}
st_linestring(matrix(10:1,5))
```


## Useful spatial operations

```{r}
# Convert to standard feature
st_as_sf()
```


```{r}
# Transform Coordinate Reference System
nc.web_mercator <- st_transform(nc, 3857)
#st_crs(nc) <- "+proj=longlat +datum=NAD27 +no_defs +ellps=clrk66 +nadgrids=@conus,@alaska,@ntv2_0.gsb,@ntv1_can.dat"
```

`st_is_valid` and `st_is_simple` return a boolean indicating whether a geometry is valid or simple.


```{r}
# Distance between geometries
st_distance(x[c(1,4,22),], x[c(1, 33,55,56),])
```

The commands `st_intersects`, `st_disjoint`, `st_touches`, `st_crosses`, `st_within`, `st_contains`, `st_overlaps`, `st_equals`, `st_covers`, `st_covered_by`, `st_equals_exact` and `st_is_within_distance` return a sparse matrix with matching (TRUE) indexes, or a full logical matrix.

The commands `st_buffer`, `st_boundary`, `st_convexhull`, `st_union_cascaded`, `st_simplify`, `st_triangulate`, `st_polygonize`, `st_centroid`, `st_segmentize`, and `st_union` return new geometries.

Commands `st_intersection`, `st_union`, `st_difference`, `st_sym_difference` return new geometries that are a function of pairs of geometries.


